---
title: 'Lab Week 5: time series and a map!'
author: "Robert Saldivar"
date: "February 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= F, warning =F)
```

## Attach packages
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(sf)
library(tmap)
library(mapview)
#library(forecast)
#library(paletteer)


```


## Monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>%
  clean_names()
```

- Make description all lower case
- only keep observations where "description" variables contain "consumption"
- remove any observations where "description" variable contains "total"

```{r}
renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>%
  filter(str_detect(description, pattern = "consumption")) %>%
  filter(!str_detect(description, pattern = "total"))
```

### Convert 'yyyymm' column to a date
```{r}
renew_date <- renew_clean %>%
  mutate(yr_mo_date = lubridate::parse_date_time(yyyymm, "ym")) %>%
  mutate(month_sep = yearmonth(yr_mo_date)) %>%
  mutate(value = as.numeric(value)) %>%
  drop_na(month_sep, value)

#Make a version where I have the month and year in separate columns
renew_parsed <- renew_date %>%
  mutate(month = month(yr_mo_date, label = T)) %>%
  mutate(year = year(yr_mo_date))
```

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep,
                                          y = value,
                                          group = description)) +
  geom_line(aes(color = description))

renew_gg
```

Updating colors with paletteer palettes:
```{r}
renew_gg +
  RColorBrewer::brewer.pal() #this code is for me to try to use r color brew because paletteer is not working for me
  scale_color_paletteer_d("calecopal::figmtn")
```

### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our ts data in a couple different ways:
```{r}

renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
#renw_ts %>% gg_season(value)

ggplot(data= renew_parsed, aes(x = month, y = value, group = year)) + 
  geom_line(aes(color= year)) +
  facet_wrap(~ description,
             ncol = 1,
             scales = "free",
             strip.position = "right")

```

### Just look at the hydroelectric energy consumption
```{r}
hydro_ts <- renew_ts %>%
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
#hydro_ts %>% gg_season(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))

```

### What if I want the quarterly average consumption for hydro?
```{r}
hydro_quarterly <- hydro_ts %>%
  index_by(year_qu = ~(yearquarter(.))) %>%
  summarise(
    ave_consumption = mean(value)
  )

head(hydro_quarterly)
```

### Decompose that hydro_ts
```{r}
dcmp <- hydro_ts %>%
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()
  
```

```{r}
hist(components(dcmp)$remainder)
```

Now lets look at ACF
```{r}
hydro_ts %>%
  ACF(value) %>%
  autoplot()
```

# Danger Danger
```{r}
hydro_model <- hydro_ts %>%
  model(
    ARIMA(value),
    ETS(value)
  ) %>%
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### Make a world map!

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

